---
title: "Detecting Unusual Polling-Table Results with MultFourier"
author: "Charles Thraves, Pablo Rivas, Felipe Subiabre"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Detecting Unusual Polling-Table Results with MultFourier}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.width = 8,
  fig.height = 4.8
)
```

# Motivation

This vignette illustrates how `MultFourier` can be used to identify unusual
outcomes in a collection of multinomial observations. The application analyzes
results from polling tables belonging to the same polling location in the 2024
Chilean municipal election.

Within a polling location, registered voters are assigned to polling tables
approximately according to alphabetical order. If voter surname is unrelated to
political preference, the underlying voting distribution should be broadly
similar across tables in the same location. This structure provides a natural
multinomial goodness-of-fit setting: we can estimate a reference probability
vector from the remaining tables in the location, and then test whether the
result observed at a given table is statistically compatible with that null
distribution.

# Data: Escuela Centenario, Iquique

The example focuses on the **Escuela Centenario** polling location in Iquique.
This location comprises 15 polling tables, numbered 276 through 290. Most tables
contain 400 registered voters, while the final two tables are smaller.

The vote matrix below records abstention, votes for four mayoral candidates,
and null or blank votes. Columns correspond to individual polling tables.

```{r data}
library(MultFourier)

votes <- matrix(
  c(
    # Abstention
     96, 101, 173,  94, 106,  88,  96, 104,  88,  98, 109,  92, 104, 57, 50,
    # Candidate A
     37,  34,  31,  34,  37,  37,  35,  32,  40,  37,  37,  38,  36, 22, 25,
    # Candidate B
     39,  37,  36,  36,  26,  29,  28,  39,  39,  34,  28,  32,  35, 15, 27,
    # Candidate C
     99, 107, 110, 108, 127, 107, 108, 125,  89, 100, 109, 103,  98, 74, 70,
    # Null and blank votes
     46,  37,  42,  41,  34,  37,  45,  26,  62,  45,  48,  49,  41, 21, 21,
    # Candidate D
     83,  84,   8,  87,  70, 102,  88,  74,  82,  86,  69,  86,  86, 49, 44
  ),
  nrow = 6,
  byrow = TRUE
)

rownames(votes) <- c(
  "Abstention",
  "Candidate A",
  "Candidate B",
  "Candidate C",
  "Null/blank",
  "Candidate D"
)

colnames(votes) <- as.character(276:290)

table_sizes <- colSums(votes)
table_sizes
```

The first 13 tables have exactly 400 registered voters each ($N_m = 400$),
while tables 289 and 290 have 238 and 237 voters, respectively.

# Visual Inspection

Because table sizes differ slightly, proportions provide a fairer visual
comparison than raw counts.

```{r heatmap, fig.cap = "Vote proportions by polling table at Escuela Centenario."}
prop <- sweep(votes, 2, table_sizes, "/")

heat_data <- as.data.frame(as.table(prop))
names(heat_data) <- c("Option", "Table", "Proportion")

if (requireNamespace("ggplot2", quietly = TRUE)) {
  ggplot2::ggplot(
    heat_data,
    ggplot2::aes(x = Table, y = Option, fill = Proportion)
  ) +
    ggplot2::geom_tile(color = "white", linewidth = 0.25) +
    ggplot2::scale_fill_gradient(low = "white", high = "black") +
    ggplot2::labs(
      x = "Polling table",
      y = NULL,
      fill = "Proportion"
    ) +
    ggplot2::theme_minimal(base_size = 11) +
    ggplot2::theme(
      panel.grid = ggplot2::element_blank(),
      axis.text.x = ggplot2::element_text(angle = 45, hjust = 1)
    )
} else {
  heatmap(
    prop,
    Rowv = NA,
    Colv = NA,
    scale = "none",
    margins = c(5, 10),
    xlab = "Polling table",
    ylab = ""
  )
}
```

Inspection of the heatmap highlights **Table 278**:
Candidate D received only 8 votes (2%), compared with 70–100 votes
(17%–25%) in neighboring tables, while abstention was markedly higher (173 votes).

While visually salient, graphical inspection alone does not quantify how
statistically improbable this outcome is under the multinomial model.

# Leave-One-Out Reference Distribution

Let $X_m = (X_{m1}, \ldots, X_{mK})$ denote the vector of observed counts at
table $m$, with total sample size $N_m = \sum_{k=1}^K X_{mk}$.

To evaluate whether table $m$ is compatible with the location as a whole without
allowing an anomalous table to contaminate its own reference distribution, we
estimate the null probability vector using a **leave-one-out** scheme:

\[
\widehat p_k^{(-m)}
=
\frac{\sum_{j \neq m} X_{jk}}{\sum_{j \neq m} N_j},
\qquad k = 1, \ldots, K.
\]

Under this formulation, table $m$ contributes zero weight to its own reference
vector $\widehat{\mathbf{p}}^{(-m)}$.

```{r loo-probabilities}
loo_p <- sapply(
  seq_len(ncol(votes)),
  function(m) {
    other_counts <- rowSums(votes[, -m, drop = FALSE])
    other_counts / sum(other_counts)
  }
)

rownames(loo_p) <- rownames(votes)
colnames(loo_p) <- colnames(votes)

# Reference distribution for table 278
round(loo_p[, "278"], 4)
```

# Detailed Analysis of Table 278

For table 278, the null hypothesis posits:

\[
X_{278} \sim \mathrm{Multinomial}\left(N_{278} = 400, \; \widehat{\mathbf{p}}^{(-278)}\right).
\]

We first compare observed counts with expected counts under $\widehat{\mathbf{p}}^{(-278)}$:

```{r expected-278}
comparison_278 <- data.frame(
  Option = rownames(votes),
  Observed = votes[, "278"],
  Expected = round(table_sizes["278"] * loo_p[, "278"], 1),
  Difference = round(votes[, "278"] - table_sizes["278"] * loo_p[, "278"], 1),
  row.names = NULL
)

knitr::kable(
  comparison_278,
  caption = "Observed vs. expected counts for Polling Table 278."
)
```

Now we compute the exact multinomial $p$-value using `pval_fourier()`. By default,
`MultFourier` employs the Log-Likelihood Ratio (`"llr"`) test statistic:

```{r one-pvalue}
fit_278 <- pval_fourier(
  x = votes[, "278"],
  p = loo_p[, "278"]
)

fit_278
```

The returned `multfourier` object provides direct access to all summary fields:
* `fit_278$pval`: computed $p$-value ($\approx 9.20 \times 10^{-33}$).
* `fit_278$stat`: test statistic used (`"Log-Likelihood Ratio"`).
* `fit_278$n_terms`: number of Fourier expansion terms evaluated (converged in 247 terms).
* `fit_278$time`: total execution time in seconds ($\approx 0.35$ s).
* `fit_278$status`: completion status ID (`0` indicates full convergence).

The resulting $p$-value on the order of $10^{-33}$ rigorously confirms that the
observed pattern at table 278 is extraordinarily unlikely under the null model.

# Screening All Polling Tables

We now iterate over all 15 tables in Escuela Centenario. Each table is evaluated
against its own leave-one-out reference distribution:

```{r all-tables-llr}
p_llr <- sapply(
  colnames(votes),
  function(m) {
    fit <- pval_fourier(x = votes[, m], p = loo_p[, m])
    fit$pval
  }
)

results_llr <- data.frame(
  Table = colnames(votes),
  Voters = table_sizes,
  p_value = format(p_llr, scientific = TRUE, digits = 4),
  row.names = NULL
)

knitr::kable(
  results_llr,
  caption = "Leave-one-out multinomial p-values across all tables (LLR statistic)."
)
```

We can visualize the results on a $-\log_{10}(p)$ scale, where higher values
indicate greater discrepancy with the reference distribution:

```{r pvalue-plot, fig.cap = "Leave-one-out -log10(p) across polling tables (LLR)."}
# Floor at 1e-35 to maintain a clean graphical scale for extreme tails
neg_log10_p <- -log10(pmax(p_llr, 1e-35))

plot(
  as.integer(names(p_llr)),
  neg_log10_p,
  type = "b",
  pch = 19,
  col = "black",
  xlab = "Polling table",
  ylab = expression(-log[10](p)),
  xaxt = "n",
  main = "Escuela Centenario: Anomaly Profile"
)
axis(1, at = as.integer(names(p_llr)), labels = names(p_llr))
abline(h = -log10(0.05 / 15), lty = 2, col = "red") # Bonferroni threshold
text(285, -log10(0.05 / 15) + 1.5, "Bonferroni threshold (alpha = 0.05)", col = "red", cex = 0.85)
```

Table 278 exhibits an astronomical signal ($-\log_{10}(p) \approx 32$), whereas
all other tables remain within typical sampling variability.

# Comparing Multinomial Test Statistics

`MultFourier` supports several classic statistics to partition and order the
multinomial sample space:

1. **Probability Mass Function (`"pmf"`)**: The exact multinomial point mass
   $\mathbb{P}(X = x)$. Outcomes with probability less than or equal to the
   observed outcome form the rejection region.
2. **Pearson's Chi-Square (`"chi2"`)**: Standard $\sum (O_k - E_k)^2 / E_k$ discrepancy.
3. **Log-Likelihood Ratio (`"llr"`)**: $G^2 = 2 \sum O_k \ln(O_k / E_k)$, standard for
   likelihood ratio tests.
4. **Power-Divergence Family (`"pd"`)**: Cressie & Read (1984) family parameterized
   by $\lambda \in \mathbb{R}$. We use $\lambda = 2/3$, a recommended choice
   balancing Pearson ($\lambda = 1$) and LLR ($\lambda \to 0$).

```{r compare-tests}
test_specs <- list(
  "PMF"                   = list(stat = "pmf"),
  "Pearson (Chi2)"        = list(stat = "chi2"),
  "Log-Likelihood Ratio"  = list(stat = "llr"),
  "Power Divergence (2/3)"= list(stat = "pd", lambda = 2/3)
)

p_by_test <- sapply(
  test_specs,
  function(spec) {
    sapply(
      colnames(votes),
      function(m) {
        args <- list(x = votes[, m], p = loo_p[, m], stat = spec$stat)
        if (!is.null(spec$lambda)) args$lambda <- spec$lambda
        fit <- do.call(pval_fourier, args)
        fit$pval
      }
    )
  }
)

comparison <- data.frame(
  Table = colnames(votes),
  round(p_by_test, 4),
  check.names = FALSE,
  row.names = NULL
)

knitr::kable(
  comparison,
  caption = "Leave-one-out p-values across different test statistics."
)
```

Notice that for typical tables (e.g. 276, 277, 282, 285), all statistics yield
comparable $p$-values. For table 278, every statistic produces a $p$-value
essentially equal to zero ($\le 10^{-13}$), showing that the detection of the
anomaly is completely robust to the choice of statistic.

```{r compare-tests-plot, fig.cap = "Comparison of -log10(p) across different test statistics."}
plot_data <- data.frame(
  Table   = rep(as.integer(colnames(votes)), times = ncol(p_by_test)),
  Test    = rep(colnames(p_by_test), each = nrow(p_by_test)),
  p_value = as.vector(p_by_test)
)

# Display transform floored at 1e-35 for visual clarity
plot_data$neg_log10_p <- -log10(pmax(plot_data$p_value, 1e-35))

if (requireNamespace("ggplot2", quietly = TRUE)) {
  ggplot2::ggplot(
    plot_data,
    ggplot2::aes(
      x = Table,
      y = neg_log10_p,
      color = Test,
      shape = Test,
      linetype = Test
    )
  ) +
    ggplot2::geom_line(linewidth = 0.6) +
    ggplot2::geom_point(size = 2.2) +
    ggplot2::scale_x_continuous(breaks = 276:290) +
    ggplot2::labs(
      x = "Polling table",
      y = expression(-log[10](p)),
      color = "Test Statistic",
      shape = "Test Statistic",
      linetype = "Test Statistic"
    ) +
    ggplot2::theme_minimal(base_size = 11) +
    ggplot2::theme(legend.position = "bottom")
} else {
  matplot(
    276:290,
    matrix(plot_data$neg_log10_p, ncol = ncol(p_by_test)),
    type = "b",
    pch = 1:4,
    lty = 1:4,
    xlab = "Polling table",
    ylab = expression(-log[10](p)),
    xaxt = "n"
  )
  axis(1, at = 276:290, labels = 276:290)
  legend("topright", legend = colnames(p_by_test), pch = 1:4, lty = 1:4, bty = "n")
}
```

# Methodological Scope and Limitations

When interpreting these results, two methodological aspects must be kept in mind:

1. **Plug-in estimation**: The leave-one-out probability vector $\widehat{\mathbf{p}}^{(-m)}$
   is estimated from data and then treated as known by `pval_fourier()`. The resulting
   numbers should strictly be interpreted as **plug-in goodness-of-fit $p$-values**.
2. **Nature of the anomaly**: A vanishingly small $p$-value indicates that table 278
   is highly incompatible with the assumption that all tables share a common multinomial
   distribution. It does **not** identify the cause of the discrepancy (e.g. data-entry
   error, transcription error, or genuine localized demographic differences).
3. **Multiple testing**: When screening many tables simultaneously, family-wise error
   rate or false discovery rate corrections (e.g. Bonferroni or Benjamini-Hochberg)
   should be applied.

# Summary

This example demonstrates a clean workflow for analyzing grouped multinomial counts:

1. **Visualize** proportions across groups to detect macroscopic deviations.
2. **Construct** leave-one-out reference probabilities to prevent circular influence.
3. **Compute** exact multinomial $p$-values using `pval_fourier()`.
4. **Compare** multiple statistics (`"llr"`, `"pmf"`, `"chi2"`, `"pd"`) to confirm robustness.
